---
title: "Measuring AI-Adopter Types: Comparison of Traditional vs. AI-Facilitated Procedures"
shorttitle: "AI--Adopters"
author:
  - name: Jane Doe
    corresponding: true
    affiliations:
      - name: Generic University
        department: Department of Scholarly Studies
suppress-title-page: true
bibliography: bibliography.bib
format:
  apaquarto-docx: default
editor: 
  markdown: 
    wrap: 72
execute:
  warning: false
  echo: false
  message: false
---

Abstract

The present study is an exploratory comparison of three approaches to developing items intended to assess employees’ orientations toward workplace AI. Items for Callari and Puppione (2025)’s qualitatively-derived ideal AI user types were generated through (a) ChatGPT, (b) AI-GENIE, and (c) industrial-organizational psychology subject-matter experts. Traditional (e.g., EFAs) and semantic (e.g., exploratory graph analyses) results suggest noteworthy differences in the factor structure and communities across the three item-generation procedures. 

Artificial intelligence (AI) has the potential to alter the fundamental
design and experience of work. The Job Characteristics Model [@hackman1976motivation; @hackman1975development] proposes that core job characteristics shape critical psychological states and, ultimately, important employee outcomes, and emerging research suggests that digitization may substantially alter these characteristics. For example, digital technologies may automate repetitive tasks, but they may also fragment work, limit skill use, or diminish employees’ sense of ownership over their work. @arora2025meaningful's review of meaningful work in the digital age
identifies autonomy, skill variety, feedback, task significance, and
task identity as job characteristics affected by workplace digitization.
As such, as AI’s role in the design and performance of work grows it is important to understand how employees engage with the technology. The present makes two key contributions to our understanding of AI in the workplace: Our first aim is to create a measure of type of AI adopter. Second, we accomplish to goal of scale creation via a novel comparison of three approaches to developing items, comparing traditionally-derived subject matter expert ratings, with LLM-derived procedures including prompting via ChatGPT and through a dedicated R package (AI-GENIE).

## Conceptualizing AI Use in the Workplace

Conceptualizing and measuring employee AI use is challenging. Existing
research operationalizes AI use in terms of the extent or intensity it
is used to perform work. For example, @zhang2026artificial measure AI
usage with items assessing the degree to which employees use AI in
carrying out their job functions. Increasing AI integration may alter
employees’ routines, role boundaries, autonomy, and skills, suggesting
that the potential impact of AI use likely goes beyond frequency or
intensity. Two employees could report similarly high levels of AI use
while interacting with the technology in fundamentally different ways.
One employee may primarily use AI to automate routine administrative
work, another may use it to generate and refine ideas, and still another
may engage with AI iteratively as a source of expertise or cognitive
challenge. Accordingly, understanding AI-enabled work may require
measures that capture how they incorporate AI into their work.  

@callari2025meaningful provide one framework for distinguish forms of AI
use. Drawing on qualitative data from employees using Microsoft 365
Copilot, they identified three ideal types of AI users
Efficiency-Seeking, Pragmatic Integrators, and Collaborative Optimizers.
These ideal types imply qualitatively different relationships between AI
and features of work including autonomy, skill utilization, cognitive
effort, and decision-making. The @callari2025meaningful framework
suggests that AI use is not a uniform behavior but rather is best
considered multidimensional.Yet, translating qualitative ideal types
into quantitative measures presents a challenge. Ideal types represent
recurring patterns of behavior and experience, but they do not
necessarily correspond to distinct psychological constructs. Additional
work is needed to determine whether AI user types such as those
presented by @callari2025meaningful can be measured with survey
items. Interestingly, the measurement of AI use is potentially enriched
by recent technological developments. For a long time, the scale
development process has relied on the efforts of researchers and
subject-matter experts to turn construct definitions to survey items
measuring the construct. Now, generative AI now provides tools for
producing surveys that measure constructs from idea generation to item
testing. However, the question of whether survey items produced by
generative AI can properly measure constructs. 

## The Current Study

The present study is an exploratory comparison of three approaches to developing items intended
to assess employees’ orientations toward workplace AI use. Specifically,
we compare items generated through (a) ChatGPT based on @callari2025meaningful’s (2025) ideal AI user types, (b) AI-GENIE, and (c)
industrial-organizational psychology subject-matter experts. We evaluate
the internal consistency and underlying factor structure of each item
set. In doing so, this study has two primary objectives: first, to
provide evidence regarding whether qualitatively derived forms of
human–AI interaction can be operationalized as quantitative constructs,
and second, to examine whether different approaches to item generation
yield comparable representations of those constructs. More broadly,
establishing how employees differ in their use of AI represents an
important step toward understanding how AI may reshape established
features of job design and the psychological experience of work.

>*Research Question 1*: Is one instrument development procedure (ChatGPT, AI-GENIE, or SMEs) yet a more coherent structure through a traditional lens?
 
>*Research Question 2*: Is one instrument development procedure (ChatGPT, AI-GENIE, or SMEs) yet a more coherent structure through a contemporary lens?

>*Research Question 3*: Does the combination of traditional and contemporary analyses provide a richer understanding of the structure than either alone?


```{r}
names <- read.csv("data/AI+Adoption_partialdata.csv")[1,]
data <- read.csv("data/AI+Adoption_partialdata.csv")[-c(1:4),]  ## subset rows when get responses

recode_map <- c(
  "Strongly disagree"          = 1,
  "Somewhat disagree"          = 2,
  "Neither agree nor disagree" = 3,
  "Somewhat agree"             = 4,
  "Strongly agree"             = 5
)

data[20:85] <- lapply(data[20:85], function(x) unname(recode_map[x]))

#Remove incomplete data rows
data$missing <- rowSums(is.na(data))
data <- data[data$missing < 25,]

################### ALPHAs
## Chat GPT

es_gpt_alpha <- psych::alpha(data[20:27], warnings=FALSE)    ## ES
pi_gpt_alpha <- psych::alpha(data[28:35])    ## PI
co_gpt_alpha <- psych::alpha(data[36:43])    ## CO

## Humans

#reflecting human-generated items 3-7
data$SME_ES_3 <- 6-data$SME_ES_3
data$SME_ES_4 <- 6-data$SME_ES_4
data$SME_ES_5 <- 6-data$SME_ES_5
data$SME_ES_6 <- 6-data$SME_ES_6 #empirical decision
data$SME_ES_7 <- 6-data$SME_ES_7

es_hum_alpha <- psych::alpha(data[44:50])    ## ES
pi_hum_alpha <- psych::alpha(data[51:57])    ## PI
co_hum_alpha <- psych::alpha(data[58:64])    ## CO

## AI Genie

es_genie_alpha <- psych::alpha(data[65:71])    ## ES
pi_genie_alpha <- psych::alpha(data[72:78])    ## PI
co_genie_alpha <- psych::alpha(data[79:85])    ## CO

##################### Demographics

referent <- table(data$Q51)  ## work or school?

data$Q55 <- as.numeric(as.character(data$Q55))  ## age
```

```{r}
#| label: tbl-alphas
#| tbl-cap: Alpha estimates all 9 subscales

library(flextable)
library(tidyverse)

tibble(
  Scale = c("Chat GPT", "", "", "SME", "", "", "AI-GENIE", "", ""),
  Subscale = c("Effeciency Seeking", "Pragmatic Integrator", "Collaborative Optimiser", "Effeciency Seeking", "Pragmatic Integrator", "Collaborative Optimiser", "Effeciency Seeking", "Pragmatic Integrator", "Collaborative Optimiser"),
  Alpha = c(es_gpt_alpha$total$raw_alpha, 
            pi_gpt_alpha$total$raw_alpha, 
            co_gpt_alpha$total$raw_alpha, 
            es_hum_alpha$total$raw_alpha, 
            pi_hum_alpha$total$raw_alpha,
            co_hum_alpha$total$raw_alpha,
            es_genie_alpha$total$raw_alpha,
            pi_genie_alpha$total$raw_alpha,
            co_genie_alpha$total$raw_alpha)          
) |>
  flextable() |>
  theme_apa() |>
  width(width = c(2.5, 2, 2)) |>
  align(align="left", part="all")

```

```{r}

library(expss)

majors <- data %>%
  tab_cells(mrset_f(X5_)) %>%
  tab_stat_cases(label = "n") %>%
  tab_stat_cpct(label = "%") %>%
  tab_pivot(stat_position = "inside_columns") %>%
  where(!grepl("#", row_labels))

major <- as_tibble(majors[-1])
names(major)[1] <- "Major"
names(major)[2] <- "Total (n)"
names(major)[3] <- "Percent"

major$`Total (n)` <- as.integer(major$`Total (n)`)

#major|>                            ## Nice table (hidden from SIOP render)
#  flextable() |>
#  theme_apa() |>
#  width(width = c(4, 2, 2)) |>
#  align(align="left", part="all")

largest <- major %>%
  arrange(`Total (n)`)

  
```

```{r}
#| tbl-cap: Means, standard deviations, and correlations
#| label: tbl-corrs

## Chat GPT

data$es_gpt <- rowMeans(data[20:27],na.rm=TRUE)    ## ES
data$pi_gpt <- rowMeans(data[28:35],na.rm=TRUE)    ## PI
data$co_gpt <- rowMeans(data[36:43],na.rm=TRUE)    ## CO

## Humans

data$es_hum <- rowMeans(data[44:50],na.rm=TRUE)    ## ES
data$pi_hum <- rowMeans(data[51:57],na.rm=TRUE)    ## PI
data$co_hum <- rowMeans(data[58:64],na.rm=TRUE)    ## CO

## AI Genie

data$es_genie <- rowMeans(data[65:71],na.rm=TRUE)    ## ES
data$pi_genie <- rowMeans(data[72:78],na.rm=TRUE)    ## PI
data$co_genie <- rowMeans(data[79:85],na.rm=TRUE)    ## CO


# Select variables
corr_vars <- data[, c(
  "es_gpt",
  "es_hum",
  "es_genie",
  "pi_gpt",
  "pi_hum",
  "pi_genie",
  "co_gpt",
  "co_hum",
  "co_genie"
)]

# Apply manuscript-friendly labels
names(corr_vars) <- c(
  "GPT (ES)",
  "SME (ES)",
  "AI-GENIE (ES)",
  "GPT (PI)",
  "SME (PI)",
  "AI-GENIE (PI)",
  "GPT (CO)",
  "SME (CO)",
  "AI-GENIE (CO)"
)


r <- corx::corx(corr_vars,                    
                triangle = "lower",
##                stars = c(0.05, 0.01, 0.001),
                describe = c(`$M$` = mean, `$SD$` = sd))

papaja::apa_table(r$apa, # apa contains the data.frame needed for apa_table
                  landscape = TRUE,      ## doesn't work
                  note="* p < 0.05; ** p < 0.01; *** p < 0.001",
                  escape = FALSE,
                  font_size = 'scriptsize')             ## this may not work if knit to Word

```
# Method

## Participants

`r nrow(data)` undergraduate students responded (Mean age = `r round(mean(data$Q55, na.rm=TRUE),2)`, SD = `r round(sd(data$Q55, na.rm=TRUE),2)`). The students were given the choice of 50 different majors to choose from, with the greatest representation being `r largest[nrow(largest),1]` at `r round(largest[nrow(largest),3],2)`%, with a second largest group of `r largest[(nrow(largest)-1),1]` at `r round(largest[(nrow(largest)-1),3],2)`%. When asked, "Which context did you have in mind as you responded to the statements in this survey?", `r round((referent[3]/nrow(data))*100,2)`% indicated "school", `r round((referent[4]/nrow(data))*100,2)`% indicated "work", and `r round((referent[2]/nrow(data))*100,2)`% stated "other". 

## Materials

Items were generated based on the three themes uncovered by @callari2025meaningful in their qualitative study on the way employees perceive meaningful work in tasks co-generated by Microsoft 365 Copilot specifically. They identified three ideal types, including an Efficiency-Seeking Type (perceives the AI tool as a straightforward task-assistance tool), a Pragmatic Integrator Type (perceives AI tool as a smarter assistant), and a Collaborative Optimiser Type (perceives AI tool an expert-like teammate). Three different methods of item generation were used (and compared) here. 
	
### Expert--Generated Items. 

Three of the researchers independently generated 6-8 items per dimension. Combined lists of items were discussed by dimension, and the initial list of 57 candidate items was reduced to *k*=18 through consensus evaluation of item clarity and content domain coverage.

### ChatGPT-Generated Items. 

The following prompt was input into ChatGPT version: “I need your help with generating a tentative list of survey items. I am reading the attached paper for the third time. I am interested in the three AI user ideal types and the survey items that could be generated to measure an employee's propensity toward one type or another.” Eight items were generated per dimension, along with rationale for those item choices (total *k*=24). See Appendix B for the full ChatGPT prompt and output, and Appendix C for a full list of items.

### AI-GENIE-Generated Items. 

AI-GENIE is a tool for automated item development and validation using large language models and network psychometrics [@russell2026generative]. The following initial prompt was input, "You are an expert scale developer and methodologist who writes high quality and succinct (although with a use a 7th grade reading level) items based on the overarching definition of AI Adopter Type as articulated in @callari2025meaningful: “Three Ideal Types were identified: Ideal Type (1) – the Efficiency-Seeking Type – perceives AI as a straightforward task-assistance tool, Ideal Type (2) – the Pragmatic Integrator Type – views AI as a smarter assistant, and Ideal Type (3) – the Collaborative Optimiser Type – considers AI an expert-like teammate." Definitions of all three of the above dimensions, as well as example descriptors are provided as part of the prompt. Thirty items were initially generated. Combined lists of items were discussed by the researchers by dimension, and the most unique and clearly-written items were retained for the survey (total *k*=21). See the Appendix for all study items.

## Procedure

After consent, students were asked to think of the context where they use AI the most frequently (for example, as a student or as an employee) and to keep that role in mind for the entire survey. Respondents were shown items within generative grouping, followed by a few demographic questions. They responded to all AI--Adopter items on a 5-point Likert-type scale ranging from 1 (*strongly agree*) to 5 (*strongly disagree*). Following the survey, respondents were debriefed and offered course credit for their time.

# Results

## Traditional Lens

Inter--scale correlations as well as scale means and standard deviations are presented in @tbl-corrs [note that the sequence of scale orderings is informed by @campbell1959convergent]. Here, the Efficiency Seeking scales exhibit the least intra--dimension agreement (correlations across development methodologies ranging from $r_{GPT-SME}$ = .14, ns to $r_{GPT-Genie}$ = .51, *p* < .001). Pragmatic Integrator and Collaborative Optimiser scales exhibited consistently higher intra--dimension associations (ranging from a low of $r_{Genie(CO)-SME(CO)}$ = .76, *p* < .001 to a high of $r_{Genie(PI)-SME(PI)}$ = .84, *p* < .001). *Lack* of association with "other" scale dimensions was most prominent with the Efficiency Seeking scales, particularly so for the ChatGPT developed scale (and to a lesser degree, the SME--derived scale; e.g., the average magnitude of correlation was *smallest*). 

Individual scale alphas were predominantly within the .8 to .9 range, although two scales exhibited suppressed values (the ChatGPT Efficiency Seeking scale $\alpha$ = .69 and the SME Efficiency Seeking scale $\alpha$ = .59). Note that: 1) these are the two scales that also exhibited the lowest average inter--scale associations (see @tbl-corrs), and 2) the numbers of items differed across scales (*k~ChatGPT~* = 6, *k~SME~* = 7, and *k~AI--Genie~* = 8). All scale $\alpha$'s are presented in @tbl-alphas

```{r efa-network}
library(tidygraph)
library(ggraph)

## EFA loading network: items link to factors with |loading| >= cutoff;
## edge width/opacity = |loading|, dashed = negative, node colour = intended subscale
efa_network <- function(items, nfactors = 3, cutoff = .30, title = NULL, seed = 2026,
                        factor_prefix = NULL) {
  efa <- psych::fa(items, nfactors = nfactors, fm = "ml", rotate = "oblimin")  # oblimin (gamma = 0) = Mplus quartimin
  L <- unclass(efa$loadings)
  if (!is.null(factor_prefix)) colnames(L) <- paste0(factor_prefix, seq_len(ncol(L)))
  factors <- colnames(L)

  edges <- as.data.frame(as.table(L), stringsAsFactors = FALSE) |>
    setNames(c("from", "to", "loading")) |>
    dplyr::filter(abs(loading) >= cutoff) |>
    dplyr::mutate(strength = abs(loading),
                  sign = ifelse(loading < 0, "Negative", "Positive"))

  subscale <- dplyr::case_when(
    grepl("_(ES|EF)_", rownames(L)) ~ "Efficiency Seeking",
    grepl("_PI_", rownames(L))      ~ "Pragmatic Integrator",
    grepl("_CO_", rownames(L))      ~ "Collaborative Optimiser"
  )
  nodes <- rbind(
    data.frame(name = factors, type = "factor", Subscale = NA),
    data.frame(name = rownames(L), type = "item", Subscale = subscale)
  )
  nodes$Subscale <- factor(nodes$Subscale,
                           levels = c("Efficiency Seeking", "Pragmatic Integrator", "Collaborative Optimiser"))

  g <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE)

  set.seed(seed)
  lay <- if (nfactors == 1) create_layout(g, layout = "star", center = 1) else create_layout(g, layout = "stress")  # star keeps 1-factor spokes from overlapping
  ggraph(lay) +
    geom_edge_link(aes(edge_width = strength, edge_alpha = strength, edge_linetype = sign),
                   colour = "grey55") +
    geom_node_point(data = \(d) dplyr::filter(d, type == "factor"),
                    colour = "grey80", size = 7) +
    geom_node_point(data = \(d) dplyr::filter(d, type == "item"),
                    aes(colour = Subscale), size = 5) +
    geom_node_text(aes(label = name), size = 3, repel = TRUE) +
    scale_edge_width(range = c(0.3, 2.5), limits = c(cutoff, 1), name = "|Loading|") +
    scale_edge_alpha(range = c(0.3, 1), limits = c(cutoff, 1), guide = "none") +
    scale_edge_linetype_manual(values = c(Positive = "solid", Negative = "dashed"),
                               name = "Direction") +
    scale_colour_brewer(palette = "Set2", drop = FALSE) +
    labs(title = title, colour = "Subscale") +
    theme_graph(base_family = "sans") +
    theme(legend.position = "bottom", legend.box = "vertical",
          plot.title = element_text(size = 14, face = "plain"))
}
```

```{r}
#| label: fig-net-gpt
#| fig-cap: EFA loading network, ChatGPT--generated items (3 factors, ML, oblimin; |loading| ≥ .30)
#| fig-width: 8
#| fig-height: 8

efa_network(data[20:43], title = "ChatGPT Items")
```

```{r}
#| label: fig-net-sme
#| fig-cap: EFA loading network, expert--generated items (3 factors, ML, oblimin; |loading| ≥ .30)
#| fig-width: 8
#| fig-height: 8

efa_network(data[44:64], title = "Expert--Generated Items")
```

```{r}
#| label: fig-net-genie
#| fig-cap: EFA loading network, AI-GENIE-generated items (3 factors, ML, oblimin; |loading| ≥ .30)
#| fig-width: 8
#| fig-height: 8

efa_network(data[65:85], title = "AI--GENIE Items")
```

Exploratory factor analyses (EFAs) were also conducted. One--, two--, and three--factor solutions were estimated for each item set using maximum likelihood estimation and an oblique Quartimin rotation. Following @hu1999cutoff's guidelines, model fit was evaluated using multiple indicators. The results presented here focus only on the three--factor models.

The three-factor model for the ChatGPT-Generated items returned the weakest structural evidence (χ²(207) = 347.09, *p* \< .001, RMSEA = .084, 90% CI \[.068, .099\], CFI = .879, TLI = .838, and SRMR = .053). Factors 1 and 2 were significantly correlated (r** = .48, *p* \< .05), while Factor 3 was did not have a significant correlation with Factor 1 (*r* = -.03) or Factor 2 (*r* = -.15). Visual representation of loadings is presented in @fig-net-gpt. 

The three-factor solution for the SME-generated items demonstrated the
strongest overall model fit of the three item-generation approaches,
χ²(150) = 183.17, *p* = .034, RMSEA = .048, 90% CI \[.015, .071\], CFI =
.962, TLI = .947, and SRMR = .050. All three factors were significantly correlated at the 5% level threshold (*p* \< .05) . Factor 1 correlated .44 with Factor 2 and .64 with Factor 3, while Factors 2 and 3 correlated .46. Overall, the three-factor model provided strong fit, although the like the AI--GENIE items, loadings did not create three distinct item groupings (see @fig-net-sme).

The AI-GENIE-generated items demonstrated fair fit to the three--factor solution, χ²(150) = 207.64, *p* = .001, RMSEA = .064, 90% CI \[.041, .084\], CFI = .960, TLI = .944, and SRMR = .036. Item loading patterns provided partial support, with several non--hypothesized associations and substantive cross--loadings (see @fig-net-genie for a visual representation)^[Each figure shows the results of an exploratory factor analysis as a network. The large grey nodes are the extracted factors. The smaller colored nodes are the survey items, and their color shows which subscale the item was written for: Efficiency Seeking, Pragmatic Integrator or Collaborative Optimiser. A line between an item and a factor means the item loads on that factor at |.30| or higher. Thicker, darker lines mean stronger loadings, and red lines mean negative loadings. If the items worked as intended, each factor would be surrounded by a cluster of items of a single color, and each item would connect to only one factor. Factors that collect items of mixed colors, or items that connect to several factors (cross-loadings), suggest the intended three-type structure is not holding up.]. Factors 1 and 2 exhibited a significant correlation (r = .76, *p* \< .05), whereas the correlations of Factor 3 with Factors 1 (*r* = .30) and 2 (*r* = .29) did not approach statistical significance. Overall, the three-factor model demonstrated relatively good fit, the loading pattern suggests two somewhat clear factors with item cross loadings and an ambiguous third factor.

Taken together, these results point toward the ChatGPT--generated items as exhibiting the weakest traditional (empirical) evidence, with the observed structure largely not conforming to the theoretical associations. Additionally, the ChatGPT-generated items demonstrated worse model fit than the corresponding SME and AI--GENIE item sets.

## Contemporary Lens

```{r}
#| label: fig-ega-gpt
#| fig-cap: Exploratory Graph Analysis (ChatGPT)

if (!requireNamespace("EGAnet", quietly = TRUE)) install.packages("EGAnet")
library(EGAnet)

ega_gpt   <- EGA(data[20:43], model = "glasso", algorithm = "walktrap", plot.EGA = FALSE)
#ega_sme   <- EGA(data[44:64], model = "glasso", algorithm = "walktrap", plot.EGA = FALSE)
#ega_genie <- EGA(data[65:85], model = "glasso", algorithm = "walktrap", plot.EGA = FALSE)

plot(ega_gpt)   + ggplot2::labs(title = "ChatGPT Items")
#plot(ega_sme)   + ggplot2::labs(title = "Expert-Generated Items")
#plot(ega_genie) + ggplot2::labs(title = "AI-GENIE Items")
```

```{r}
#| label: fig-ega-sme
#| fig-cap: Exploratory Graph Analysis (SME)

if (!requireNamespace("EGAnet", quietly = TRUE)) install.packages("EGAnet")
library(EGAnet)

#ega_gpt   <- EGA(data[20:43], model = "glasso", algorithm = "walktrap", plot.EGA = FALSE)
ega_sme   <- EGA(data[44:64], model = "glasso", algorithm = "walktrap", plot.EGA = FALSE)
#ega_genie <- EGA(data[65:85], model = "glasso", algorithm = "walktrap", plot.EGA = FALSE)

#plot(ega_gpt)   + ggplot2::labs(title = "ChatGPT Items")
plot(ega_sme)   + ggplot2::labs(title = "Expert-Generated Items")
#plot(ega_genie) + ggplot2::labs(title = "AI-GENIE Items")
```

```{r}
#| label: fig-ega-genie
#| fig-cap: Exploratory Graph Analysis (AI-GENIE)

if (!requireNamespace("EGAnet", quietly = TRUE)) install.packages("EGAnet")
library(EGAnet)

#ega_gpt   <- EGA(data[20:43], model = "glasso", algorithm = "walktrap", plot.EGA = FALSE)
#ega_sme   <- EGA(data[44:64], model = "glasso", algorithm = "walktrap", plot.EGA = FALSE)
ega_genie <- EGA(data[65:85], model = "glasso", algorithm = "walktrap", plot.EGA = FALSE)

#plot(ega_gpt)   + ggplot2::labs(title = "ChatGPT Items")
#plot(ega_sme)   + ggplot2::labs(title = "Expert-Generated Items")
plot(ega_genie) + ggplot2::labs(title = "AI-GENIE Items")
```

Network Psychometrics has provided not merely a newer philosophical lens on psychological measurement, but it has also introduced novel empirical procedures [@epskamp2018network; @borsboom2022possible; @borsboom2013network]. 

Exploratory graph analysis (EGA) is a procedure that, similar to traditional factor analysis, is aimed at discovering how items on a psychological measure group together. Each item is treated as a node, and the strength of the unique association between items (after accounting for all other items) is shown as an edge; items that remain strongly linked tend to form clusters. A community-detection algorithm then identifies those clusters, which are interpreted as the dimensions or factors of the construct. An EGA plot therefore shows both the estimated structure and the item membership of each dimension in one figure.

Exploratory Graph Analyses do not specify factors in advance. Instead they estimates a network directly among items via semantic association, and the results of these are presented in @fig-ega-gpt, @fig-ega-sme, and @fig-ega-genie. Each line is a regularized partial correlation (EBICglasso), meaning the relationship between two items after controlling for all the other items. Thicker lines are stronger associations, and the line colour shows whether the association is positive or negative. Weak relationships are shrunk to zero and dropped. The Walktrap algorithm then detects communities: groups of items that are more tightly connected to each other than to the rest of the network. Unlike in the EFA plots, the node colours here mark these empirically detected communities, not the intended subscales. The number of communities is the EGA estimate of the dimensionality of each item set. 

# Discussion

The purpose of this stuy was twofold: 1) to create a measure of type of AI adopter, 2) to compare three approaches to developing items, comparing traditionally-derived subject matter expert ratings, with LLM-derivded procedures using ChatGPT and AI-GENIE. The primary question of interest was whether the AI-driven item-generation procedures (i.e., ChatGPT or AI-GENIE) as good as, or even superior to, traditionally generated SME items. 

First, all three procedures did result in a measure of AI adopter type. The results of the traditional analyses suggest the the SME-generated items demonstrated the best fit, followed by AI-GENIE and then ChatGPT. However, items did not load cleanly as expected within dimensions. The semantic exploratory graph analyses...

>DISCUSSION CENTER: Resolving the EFA visuals with the EGA visuals -- also discussing the possibility that there are not "three dimensions" but rather three different levels of AI comfort: introductory, intermediate, and expert... it's a matter of degree not type, and therefore the measurement results from three different methodologies may reasonably be expected to not differentiate greatly (@tbl-corrs), or to result in muddled structures when imposed on the data (EFAs). Implications are that this construct may benefit from further THEORETICAL specification

## Implications

One of the theoretical contributions we make is to explore the ways AI is being used at present. @callari2025meaningful presented three ideal types of AI users that we based our quantitative scale creation on. The somewhat mixed results suggest that there may not be distinct types of AI users, but rather, we can conceptualize AI use on an continuum of expertise, whereby greater levels of expertise are evidenced by a higher degree of AI collaboration. 
  
## Limitations & Future Directions

We note a number of limitations that future study/comparisons would benefit from. First, for this SIOP submission, we present a modest sample size. An additional study comparing these three procedures with a larger sample size is warranted. In addition, this first pilot sample was obtained from university students were were asked to keep one context in mind as they responded to the survey. The majority of respondents indicated they had school in mind, whereas a smaller percentage had work in mind. Future research should incorporate full time employees across industries, and account for degree of AI integration within their workplaces. 

## Conclusion

Please add after viewing 

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

