---
title: "Placeholder"
format: html
execute:
  echo: false
  warning: false
  message: false
---

```{r}
names <- read.csv("data/AI+Adoption_partialdata.csv")[1,]
data <- read.csv("data/AI+Adoption_partialdata.csv")[-c(1:4),]  ## subset rows when get responses

recode_map <- c(
  "Strongly disagree"          = 1,
  "Somewhat disagree"          = 2,
  "Neither agree nor disagree" = 3,
  "Somewhat agree"             = 4,
  "Strongly agree"             = 5
)

data[20:85] <- lapply(data[20:85], function(x) unname(recode_map[x]))

#Remove incomplete data rows
data$missing <- rowSums(is.na(data))
data <- data[data$missing < 25,]

################### ALPHAs
## Chat GPT

es_gpt_alpha <- psych::alpha(data[20:27])    ## ES
pi_gpt_alpha <- psych::alpha(data[28:35])    ## PI
co_gpt_alpha <- psych::alpha(data[36:43])    ## CO

## Humans

#reflecting human-generated items 3-7
data$SME_ES_3 <- 6-data$SME_ES_3
data$SME_ES_4 <- 6-data$SME_ES_4
data$SME_ES_5 <- 6-data$SME_ES_5
data$SME_ES_6 <- 6-data$SME_ES_6 #empirical decision
data$SME_ES_7 <- 6-data$SME_ES_7

es_hum_alpha <- psych::alpha(data[44:50])    ## ES
pi_hum_alpha <- psych::alpha(data[51:57])    ## PI
co_hum_alpha <- psych::alpha(data[58:64])    ## CO

## AI Genie

es_genie_alpha <- psych::alpha(data[65:71])    ## ES
pi_genie_alpha <- psych::alpha(data[72:78])    ## PI
co_genie_alpha <- psych::alpha(data[79:85])    ## CO

##################### Demographics

referent <- table(data$Q51)  ## work or school?

data$Q55 <- as.numeric(as.character(data$Q55))  ## age
```

## Method

## Participants

XX undergraduate and graduate students from a mid-sized midwestern university completed a survey for course credit during September of 2026. XX different majors were represented, with the highest percentages coming from Business Administration  (n = ), Construction  (n = ), and Psychology (n = ). The mean age of respondents was XX (SD = XX), and the breakdown for gender was as follows: XX% men, XX% women, XX% XX% middle gender, XX% self-described, and XX% preferred not to report their gender. The majority of the student reported being employed (XX%). 

## Materials

	Items were generated based on the three themes uncovered by Callari and Puppione (2025) in their qualitative study on the way employees perceive meaningful work in tasks co-generated by Microsoft 365 Copilot specifically. They identified three ideal types, including an Efficiency-Seeking Type (perceives the AI tool as a straightforward task-assistance tool), a Pragmatic Integrator Type (perceives AI tool as a smarter assistant), and a Collaborative Optimiser Type (perceives AI tool an expert-like teammate). Three different methods of item generation were used (and compared) here. 
	
### Expert-Generated AI Adopter Type Items. 

Three of the researchers independently generated 6-8 items per dimension. Combined lists of items were discussed by dimension, and the most unique and clearly-written items were retained for the survey. See Appendix A.

### ChatGPT-Generated AI Adopter Type Items. 

The following prompt was input into ChatGPT version: “I need your help with generating a tentative list of survey items. I am reading the attached paper for the third time. I am interested in the three AI user ideal types and the survey items that could be generated to measure an employee's propensity toward one type or another.” Eight items were generated per dimension, along with rationale for those item choices. See Appendix B for the full ChatGPT prompt and output, and Appendix C for a full list of items.

### AI-GENIE-Generated AI Adopter Type Items. 

AI-GENIE is a tool for automated item development and validation using large language models and network psychometrics [@russellgenerative]. The following initial prompt was input, "You are an expert scale developer and methodologist who writes high quality and succinct (although with a use a 7th grade reading level) items based on the overarching definition of AI Adopter Type as articulated in Callari & Puppione (2025): “Three Ideal Types were identified: Ideal Type [1] – the Efficiency-Seeking Type – perceives AI as a straightforward task-assistance tool, Ideal Type (2) – the Pragmatic Integrator Type – views AI as a smarter assistant, and Ideal Type (3) – the Collaborative Optimiser Type – considers AI an expert-like teammate." Definitions of all three of the above dimensions, as well as example descriptors are provided as part of the prompt. Thirty items were initially generated. Combined lists of items were discussed by the researchers by dimension, and the most unique and clearly-written items were retained for the survey. See Appendix D for the R Script and Appendix E for the items.

### Demographics. 

Students were asked to report their major, age, gender, work status, and job title. 

## Procedure

	After consent, students asked to think of the context where they use AI the most frequently, either work or school for example, as a student or as an employee, and to keep that role in mind for the entire survey. Respondents were shown items grouped by method, followed by the demographics. They responded to the AI-Adopter items on a 5-point Likert-type scale ranging from 1 (strongly agree) to 5 (strongly disagree). Following the survey, respondents were debriefed and offered course credit for their time.

```{r alphas}
#| label: tbl-alphas
#| tbl-cap: Alpha estimates all 9 subscales

library(flextable)
library(tidyverse)

tibble(
  Scale = c("Chat GPT", "", "", "Human", "", "", "AI Genie", "", ""),
  Subscale = c("Effeciency Seeking", "Pragmatic Integrator", "Collaborative Optimiser", "Effeciency Seeking", "Pragmatic Integrator", "Collaborative Optimiser", "Effeciency Seeking", "Pragmatic Integrator", "Collaborative Optimiser"),
  Alpha = c(es_gpt_alpha$total$raw_alpha, 
            pi_gpt_alpha$total$raw_alpha, 
            co_gpt_alpha$total$raw_alpha, 
            es_hum_alpha$total$raw_alpha, 
            pi_hum_alpha$total$raw_alpha,
            co_hum_alpha$total$raw_alpha,
            es_genie_alpha$total$raw_alpha,
            pi_genie_alpha$total$raw_alpha,
            co_genie_alpha$total$raw_alpha)          
) |>
  flextable() |>
  theme_apa() |>
  width(width = c(2.5, 2, 2)) |>
  align(align="left", part="all")

```

```{r majors}
#| label: tbl-majors
#| tbl-cap: Indicated major (current study respondents)

library(expss)

majors <- data %>%
  tab_cells(mrset_f(X5_)) %>%
  tab_stat_cases(label = "n") %>%
  tab_stat_cpct(label = "%") %>%
  tab_pivot(stat_position = "inside_columns") %>%
  where(!grepl("#", row_labels))

major <- as_tibble(majors[-1])
names(major)[1] <- "Major"
names(major)[2] <- "Total (n)"
names(major)[3] <- "Percent"

major$`Total (n)` <- as.integer(major$`Total (n)`)

major|>
  flextable() |>
  theme_apa() |>
  width(width = c(4, 2, 2)) |>
  align(align="left", part="all")

largest <- major %>%
  arrange(`Total (n)`)

#most <- largest[nrow(largest),1]
#freq <- largest[nrow(largest),3]
  
```

`r nrow(data)` undergraduate students responded (mean age = `r round(mean(data$Q55, na.rm=TRUE),2)`, sd = `r round(sd(data$Q55, na.rm=TRUE),2)`). The students were given the choice of 50 different majors to choose from, with the greatest representation being `r largest[nrow(largest),1]` at `r round(largest[nrow(largest),3],2)`% (see @tbl-majors). When asked, "Which context did you have in mind as you responded to the statements in this survey?", `r round((referent[3]/nrow(data))*100,2)`% indicated "school", `r round((referent[4]/nrow(data))*100,2)`% indicated "work", and `r round((referent[2]/nrow(data))*100,2)`% stated "other". 

Individual scale alphas are presented in @tbl-alphas

Appendix A
Expert-generated Items
Efficiency Seeking Items
1.      I am confident in my ability to construct useful AI prompts. 
2.      Using AI tools gives me more time to connect with colleagues. 
3.      AI tools often produce inaccurate results. (R)
4.      It requires a lot of effort to construct prompts that provide the results I expect. (R)
5.      AI tools cannot replace human expertise in workplace decision-making.
6.      I reserve AI for tasks that demand little mental effort.
7.      I prefer to maintain independent control over my work.
Pragmatic Integrator Items
8.      Using AI tools helps me be more creative at work. 
9.      AI tools help me to “get started” when I am stuck on a task. 
10. Over time I have gotten better at creating effective prompts for AI tools. 
11. AI offers valuable guidance on complex or unfamiliar topics.
12. My decision making is aided by AI
13. I am more productive when using AI
14. My confidence in solving a problem is enhanced with AI.
 Collaborative Optimizer Items
15. AI tools are adequate collaborators for complex tasks. 
16. My employer has done a good job with integrating AI tools into our workflows. 
17. Using AI tools involves an evolving relationship, much like getting to know a new colleague or friend.
18. Continuous workplace learning to adapt to changing skill requirements is essential in our evolving AI landscape.
19. My collaboration with AI is full of back-and-forth exchanges
20. My ability to think strategically is enhanced with AI.
21. AI challenges my thinking in beneficial, thought-provoking ways
 
Appendix B
ChatGPT Prompt
“I need your help with generating a tentative list of survey items. I am reading the attached paper for the third time. I am interested in the three AI user ideal types and the survey items that could be generated to measure an employee's propensity toward one type or another.”
Yes. Below is a tentative item pool based on Callari and Puppione’s three ideal types: Efficiency-Seeking, Pragmatic Integrator, and Collaborative Optimiser. I would treat these as source-informed draft items, not as a validated measure yet. The paper identifies these types qualitatively, not as a psychometric scale.
The article frames the three types as different ways employees conceptualize M365 Copilot: as a straightforward task-assistance tool, a smarter assistant, or an expert-like teammate. It also organizes these types around work practices, autonomy/decision-making, teamwork, and organizational support.
Suggested stem
When I use AI tools at work, I generally…
Use a 1–5 or 1–7 agreement scale.
________________________________________
AI User Ideal-Type Propensity Items
Ideal type	Draft survey item
Efficiency-Seeking Type	1. I primarily use AI tools to complete routine or administrative tasks more efficiently.
	2. I prefer to use AI tools for simple tasks rather than complex or strategic work.
	3. I see AI as a tool that helps me save time, not as a collaborator in my work.
	4. I prefer to keep final judgment and decision-making clearly in my own hands when using AI.
	5. I am cautious about relying too heavily on AI-generated outputs.
	6. I use AI most when it helps me reduce repetitive work.
	7. I prefer AI tools that require minimal prompting or adjustment from me.
	8. AI is most useful to me when it supports basic task execution without changing how I think about my work.
Pragmatic Integrator Type	1. I use AI tools as a competent assistant that helps me develop ideas or organize my thinking.
	2. I treat AI-generated outputs as a starting point that I refine using my own expertise.
	3. I use AI to help me approach tasks I might not have had enough time or knowledge to complete on my own.
	4. I see prompting AI as a useful skill that improves with practice.
	5. I use AI to support my decision-making, but I do not let it make decisions for me.
	6. AI helps me become more productive by helping me start, structure, or improve my work.
	7. I use AI to explore alternative perspectives before reaching my own conclusions.
	8. I value AI when it extends my capabilities while still leaving me responsible for the final output.
Collaborative Optimiser Type	1. I use AI as a collaborative partner that can help me think through complex issues.
	2. I value AI most when it challenges my assumptions or improves my reasoning.
	3. I see AI as a kind of expert-like teammate rather than merely a task-support tool.
	4. I use AI in an iterative back-and-forth process to improve the quality of my work.
	5. I believe effective AI use requires sophisticated prompting and reflection.
	6. AI helps me deepen my analysis rather than simply complete tasks faster.
	7. I use AI to test, refine, or strengthen my professional judgment.
	8. Working with AI can feel like collaborating with a knowledgeable partner who helps improve my thinking.
________________________________________
Why these items map onto the paper
The Efficiency-Seeking Type items reflect the paper’s description of employees who view Copilot as “just a tool” for routine tasks, time savings, and efficiency, while retaining autonomy and avoiding overdependence. The paper notes that these users value AI for automating mundane tasks and preserving human judgment.
The Pragmatic Integrator Type items reflect employees who see Copilot as a “competent assistant,” useful for initiating work, brainstorming, supporting decision-making, and extending productivity while still requiring human review and expertise.
The Collaborative Optimiser Type items reflect users who imagine AI as an expert-like teammate or “sparring partner” that can support, challenge, and refine their professional judgment. The paper emphasizes iterative interaction, prompting skill, and AI as a collaborator in knowledge work.
Recommended scoring approach
I would not force respondents into one category immediately. Instead, score each type as a separate subscale:
•	Efficiency-Seeking score = mean of Efficiency-Seeking items
•	Pragmatic Integrator score = mean of Pragmatic Integrator items
•	Collaborative Optimiser score = mean of Collaborative Optimiser items
Then you could classify respondents by their highest subscale score, or use latent profile analysis if you have enough data. This is important because the paper treats the ideal types as qualitative patterns of engagement, not as mutually exclusive validated personality categories. The authors also note that the types may reflect progressive levels of integration, but primarily capture the current adoption landscape and employees’ situated responses to AI implementation.
Shorter 12-item version
For an initial survey, I would consider starting with four items per type:
Type	Best initial items
Efficiency-Seeking	“I primarily use AI tools to complete routine or administrative tasks more efficiently.”
	“I see AI as a tool that helps me save time, not as a collaborator in my work.”
	“I prefer to keep final judgment and decision-making clearly in my own hands when using AI.”
	“I prefer AI tools that require minimal prompting or adjustment from me.”
Pragmatic Integrator	“I treat AI-generated outputs as a starting point that I refine using my own expertise.”
	“I use AI to help me develop ideas or organize my thinking.”
	“I see prompting AI as a useful skill that improves with practice.”
	“I value AI when it extends my capabilities while still leaving me responsible for the final output.”
Collaborative Optimiser	“I use AI as a collaborative partner that can help me think through complex issues.”
	“I value AI most when it challenges my assumptions or improves my reasoning.”
	“I use AI in an iterative back-and-forth process to improve the quality of my work.”
	“Working with AI can feel like collaborating with a knowledgeable partner who helps improve my thinking.”
For scale development, I would start with the longer pool, pilot it, examine factor structure, then reduce to the strongest 3–5 items per type.


“What is your rationale for generating the items you did? And why this prompt and response format?”

My rationale was to convert the paper’s qualitative ideal types into a tentative quantitative item pool while staying close to the authors’ conceptual structure. I treated the items as indicators of propensity toward an AI-use orientation, not as a validated scale or as a forced classification.
Core rationale
The paper identifies three ideal types of AI users: the Efficiency-Seeking Type, who views Copilot as a task-assistance tool; the Pragmatic Integrator Type, who views it as a smarter assistant; and the Collaborative Optimiser Type, who views it as an expert-like teammate. My items were designed to operationalize those three orientations in survey form.
I built the items around four dimensions the authors use to distinguish the types: AI-enabled work practices and adaptation, human agency/autonomy/decision-making, teamwork in human-AI environments, and organizational support. The reason the items cover more than simple “AI usefulness” is that the paper’s typology is not only about whether employees like AI. It is about how employees position AI in relation to their own expertise, autonomy, work practices, decision-making, and relationships with others.
Why those specific item themes?
For the Efficiency-Seeking Type, I emphasized routine task completion, time savings, limited prompting, caution about reliance, and preservation of human judgment. That follows the paper’s description of this type as seeing Copilot as “just a tool” that supports basic task execution, efficiency, and bounded use.
For the Pragmatic Integrator Type, I emphasized AI as a starting point, assistant, brainstorming partner, productivity enhancer, and support for decision-making while retaining human responsibility. This reflects the paper’s middle category: employees who do not merely automate routine work but use AI to extend their capability, organize thinking, and refine outputs.
For the Collaborative Optimiser Type, I emphasized AI as a partner, challenger, sparring partner, expert-like teammate, and tool for deeper analysis. This follows the paper’s description of this type as involving more iterative, cognitively demanding, and collaborative interaction with AI.
Why the prompt: “When I use AI tools at work, I generally…”
I used that prompt because it does three useful things.
First, it keeps the items behaviorally and contextually anchored. The construct is not general AI attitude; it is workplace AI-use orientation.
Second, it avoids overcommitting to Microsoft Copilot specifically. The paper is about M365 Copilot, but your future survey may involve AI tools more broadly. “AI tools at work” keeps the measure adaptable while preserving the workplace context.
Third, it frames the items as propensity statements rather than ability, frequency, or moral judgment. That matters because an employee could be capable of using AI collaboratively but still prefer to use it only for routine efficiency.
Why a Likert agreement format?
A Likert-type response format is appropriate at this early stage because you are trying to estimate degree of endorsement for each orientation. I would use something like:
1 = Strongly disagree
2 = Disagree
3 = Somewhat disagree
4 = Neither agree nor disagree
5 = Somewhat agree
6 = Agree
7 = Strongly agree
I would avoid a forced-choice format at this stage because the paper’s ideal types are analytical constructions, not necessarily mutually exclusive categories. The authors explicitly use ideal-type analysis to preserve distinctive user perspectives and capture varied patterns of AI adoption. In practice, an employee may score high on both Pragmatic Integrator and Collaborative Optimiser tendencies.
Why 8 items per type?
The 8-item-per-type structure was intentional for item development. At the early item-pool stage, I would rather overgenerate slightly than prematurely narrow the construct. The goal is to give you enough item coverage to later examine:
•	factor structure,
•	item-total correlations,
•	cross-loadings,
•	reliability,
•	discriminant validity among the three types,
•	and whether a shorter 3–5 item version per type can be retained.
So the first version is not meant to be the final scale. It is a developmental item pool.
Bottom line
The items were generated by translating the paper’s qualitative typology into survey indicators of three AI-use orientations: tool-based efficiency, assistant-based integration, and teammate-based collaboration. The prompt and Likert response format were chosen because they allow respondents to express varying degrees of alignment with each type without forcing them into a single category too early.
 
Appendix C
ChatGPT-Generated Items
Efficiency-Seeking Type
1.	I primarily use AI tools to complete routine or administrative tasks more efficiently.
2.	I prefer to use AI tools for simple tasks rather than complex or strategic work.
3.	I see AI as a tool that helps me save time, not as a collaborator in my work.
4.	I prefer to keep final judgment and decision-making clearly in my own hands when using AI.
5.	I am cautious about relying too heavily on AI-generated outputs.
6.	I use AI most when it helps me reduce repetitive work.
7.	I prefer AI tools that require minimal prompting or adjustment from me.
8.	AI is most useful to me when it supports basic task execution without changing how I think about my work.
Pragmatic Integrator Type
9.	I use AI tools as a competent assistant that helps me develop ideas or organize my thinking.
10.	I treat AI-generated outputs as a starting point that I refine using my own expertise.
11.	I use AI to help me approach tasks I might not have had enough time or knowledge to complete on my own.
12.	I see prompting AI as a useful skill that improves with practice.
13.	I use AI to support my decision-making, but I do not let it make decisions for me.
14.	AI helps me become more productive by helping me start, structure, or improve my work.
15.	I use AI to explore alternative perspectives before reaching my own conclusions.
16.	I value AI when it extends my capabilities while still leaving me responsible for the final output.
Collaborative Optimizer Type
17.	I use AI as a collaborative partner that can help me think through complex issues.
18.	I value AI most when it challenges my assumptions or improves my reasoning.
19.	I see AI as a kind of expert-like teammate rather than merely a task-support tool.
20.	I use AI in an iterative back-and-forth process to improve the quality of my work.
21.	I believe effective AI use requires sophisticated prompting and reflection.
22.	AI helps me deepen my analysis rather than simply complete tasks faster.
23.	I use AI to test, refine, or strengthen my professional judgment.
24.	Working with AI can feel like collaborating with a knowledgeable partner who helps improve my thinking.
 
Appendix D
AI-GENIE R Script
default_item_attributes_dt<- list("Efficiencyseeking"=c("summarising", "proofreading", "speed", "accuracy", "automate"),
                                  "PragmaticIntegrator"= c("assistant", "competent", "suggestions", "starter", "insights"),
                                  "CollaborativeOptimizer"= c("expert", "trusted", "complex", "smart", "understanding")
                                  )

item_type_definitions_dt<- list("Efficiencyseeking"= paste("AI is viewed as a useful tool for automating routine and mundane tasks, particularly in administrative areas. AI provides support in administrative functions, such as taking meeting minutes, summarising discussions, and proofreading emails. These features allow users to reduce the manual effort involved in repetitive tasks, enhancing the speed and accuracy of their workflows and creating a smoother operational process. AI-assisted automated tasks generally demand little mental effort and minimal oversight beyond the initial setup and occasional review of outputs. These outputs seamlessly integrate with familiar workflows, further reducing the cognitive load associated with manually performing these tasks"),
                             "PragmaticIntegrator"= paste("AI is viewed as a competent assistant, valued for its ability to generate ideas, offer insights, and support the early stages of decision-making. AI provides foundational information or preliminary suggestions, which can be refined and enhanced through user expertise and judgement. Here, work remains meaningful as employees appreciate the chance to consider alternative perspectives, with decision-making further enhanced through AI’s input. While reviewing and adjusting AI outputs requires some mental effort, users generally view this as manageable, especially compared to the more extensive work involved in initiating tasks from scratch."),
                             "CollaborativeOptimizer"= paste("Users express a desire for AI to evolve beyond a task execution tool or “junior assistant,” envisioning it as a collaborative digital “sparring partner”—akin to a trusted teammate. Users are optimistic that, as AI becomes more sophisticated, it could extend its support to more complex and strategic issues, enabling meaningful exchanges of ideas. The terminology itself underscores this perspective, framing AI as a “smart”, AI-powered expert capable of simulating a “back-and-forth” dialogue, offering alternative perspectives, challenging assumptions, and helping to identify potential areas for improvement and deeper understanding within the topic of interaction."))


item_examples_dt <- data.frame(
  type = c(
    "Efficiencyseeking",
    "Efficiencyseeking",
    "PragmaticIntegrator",
    "PragmaticIntegrator",
    "CollaborativeOptimizer",
    "CollaborativeOptimizer"
  ),

  attribute = c(
    "speed",
    "automate",
    "insights",
    "starter",
    "complex",
    "trusted"
  ),

  statement = c(
    "Using AI tools gives me more time to connect with colleagues.",
    "I reserve AI for tasks that demand little mental effort.",
    "Using AI tools helps me be more creative at work.",
    "AI tools help me to get started when I am stuck on a task.",
    "AI tools are adequate collaborators for complex tasks.",
    "Using AI tools involves an evolving relationship, much like getting to know a new colleague or friend."
  ),

  stringsAsFactors = FALSE
)

system_role_dt<- "You are an expert scale developer and methodologist who writes high quality and succinct (although with a use a 7th grade reading level) items based on the overarching definition of AI Adopter Type as articulated in Callari & Puppione (2025): “Three Ideal Types were identified:Ideal Type [1] – the Efficiency-Seeking Type – perceives AI as a straightforward task-assistance tool, Ideal Type (2) – the Pragmatic Integrator Type – views AI as a smarter assistant, and Ideal Type (3) – the Collaborative Optimiser Type – considers AI an expert-like teammate."

 
Appendix E
AI-GENIE-Generated Items
Efficiency-Seeking Items
1.	AI helps me distill lengthy discussions into concise insights.
2.	AI accelerates my task completion by automating routine processes.
3.	AI is my go-to solution for automating tedious administrative functions.
4.	AI effectively condenses complex information into actionable summaries for me.
5.	The speed of my daily operations is enhanced by AI's task automation.
6.	AI increases my confidence in the clarity and correctness of generated report summaries.
7.	AI seamlessly automates repetitive administrative chores, freeing up mental bandwidth.
Pragmatic Integrator Items
8.	AI's competent task handling enhances my workflow efficiency.
9.	AI-generated suggestions broaden the range of options I consider for creative projects.
10.	AI jumpstarts my projects by offering preliminary ideas that I can develop further.
11.	I view AI as an indispensable resource in bolstering my task management.
12.	AI systems provide reliable foundational work that I can build on.
13.	AI's recommendations offer valuable alternative pathways in my work.
14.	AI lays the groundwork for my projects, offering a springboard for development.
Collaborative Optimizer Items
15.	I perceive AI as providing an expert-level perspective in collaborative processes.
16.	In my view, AI functions as a reliable contributor, much like a dependable team member.
17.	AI thrives in environments requiring intricate problem-solving and strategic insights.
18.	AI impressively adapts to new challenges, showcasing intelligent decision-making in evolving contexts.
19.	I find AI's nuanced understanding enhances collaborative efforts by providing well-rounded viewpoints.
20.	I rely on AI to maintain reliable exchanges, much like a seasoned confidant in complex discussions.
21.	AI excels in dissecting multifaceted problems, enhancing our collaborative depth.


