---
title: "Placeholder"
format: html
execute:
  echo: false
  warning: false
  message: false
---

```{r}
names <- read.csv("data/AI+Adoption_partialdata.csv")[1,]
data <- read.csv("data/AI+Adoption_partialdata.csv")[-c(1:4),]  ## subset rows when get responses

recode_map <- c(
  "Strongly disagree"          = 1,
  "Somewhat disagree"          = 2,
  "Neither agree nor disagree" = 3,
  "Somewhat agree"             = 4,
  "Strongly agree"             = 5
)

data[20:85] <- lapply(data[20:85], function(x) unname(recode_map[x]))

################### ALPHAs
## Chat GPT

es_gpt_alpha <- psych::alpha(data[20:27])    ## ES
pi_gpt_alpha <- psych::alpha(data[28:35])    ## PI
co_gpt_alpha <- psych::alpha(data[36:43])    ## CO

## Humans

es_hum_alpha <- psych::alpha(data[44:50])    ## ES
pi_hum_alpha <- psych::alpha(data[51:57])    ## PI
co_hum_alpha <- psych::alpha(data[58:64])    ## CO

## AI Genie

es_genie_alpha <- psych::alpha(data[65:71])    ## ES
pi_genie_alpha <- psych::alpha(data[72:78])    ## PI
co_genie_alpha <- psych::alpha(data[79:85])    ## CO

##################### Demographics

referent <- table(data$Q51)  ## work or school?

data$Q55 <- as.numeric(as.character(data$Q55))  ## age
```

```{r alphas}
#| label: tbl-alphas
#| tbl-cap: Alpha estimates all 9 subscales

library(flextable)
library(tidyverse)

tibble(
  Scale = c("Chat GPT", "", "", "Human", "", "", "AI Genie", "", ""),
  Subscale = c("Effeciency Seeking", "Pragmatic Integrator", "Collaborative Optimiser", "Effeciency Seeking", "Pragmatic Integrator", "Collaborative Optimiser", "Effeciency Seeking", "Pragmatic Integrator", "Collaborative Optimiser"),
  Alpha = c(es_gpt_alpha$total$raw_alpha, 
            pi_gpt_alpha$total$raw_alpha, 
            co_gpt_alpha$total$raw_alpha, 
            es_hum_alpha$total$raw_alpha, 
            pi_hum_alpha$total$raw_alpha,
            co_hum_alpha$total$raw_alpha,
            es_genie_alpha$total$raw_alpha,
            pi_genie_alpha$total$raw_alpha,
            co_genie_alpha$total$raw_alpha)          
) |>
  flextable() |>
  theme_apa() |>
  width(width = c(2.5, 2, 2)) |>
  align(align="left", part="all")

```

```{r majors}
#| label: tbl-majors
#| tbl-cap: Indicated major (current study respondents)

library(expss)

majors <- data %>%
  tab_cells(mrset_f(X5_)) %>%
  tab_stat_cases(label = "n") %>%
  tab_stat_cpct(label = "%") %>%
  tab_pivot(stat_position = "inside_columns") %>%
  where(!grepl("#", row_labels))

major <- as_tibble(majors[-1])
names(major)[1] <- "Major"
names(major)[2] <- "Total (n)"
names(major)[3] <- "Percent"

major$`Total (n)` <- as.integer(major$`Total (n)`)

major|>
  flextable() |>
  theme_apa() |>
  width(width = c(4, 2, 2)) |>
  align(align="left", part="all")

largest <- major %>%
  arrange(`Total (n)`)

#most <- largest[nrow(largest),1]
#freq <- largest[nrow(largest),3]
  
```

`r nrow(data)` undergraduate students responded (mean age = `r round(mean(data$Q55, na.rm=TRUE),2)`, sd = `r round(sd(data$Q55, na.rm=TRUE),2)`). The students were given the choice of 50 different majors to choose from, with the greatest representation being `r largest[nrow(largest),1]` at `r largest[nrow(largest),3]`% (see @tbl-majors). When asked, "Which context did you have in mind as you responded to the statements in this survey?", `r round((referent[3]/nrow(data))*100,2)`% indicated "school", `r round((referent[4]/nrow(data))*100,2)`% indicated "work", and `r round((referent[2]/nrow(data))*100,2)`% stated "other". 

Individual scale alphas are presented in @tbl-alphas


